---
title: "Tutorial"
author: "Jack Halverson"
date: "10/11/2021"
output: html_document
---

### Load Packages
```{r}
library(leaflet)
library(sf)
library(tidyverse)
library(tidycensus)
library(ggthemes)
library(ggspatial)
library(htmlwidgets)
library(tidytransit)
```

### Get Census Data and Create New % Column
```{r}
Suffolk_Black <- get_decennial(geography = "block",
                          state = "MA", county = "Suffolk",
                          year = 2010,
                          output = "wide",
                          variables = c(tot_pop = 'P008001',
                                        bl_pop = 'P008004'),
                          geometry = TRUE)

Suffolk_Black <- Suffolk_Black %>%
  mutate(pct_Black = bl_pop / tot_pop) %>%
  filter(tot_pop > 0)
```

### Create Static Map
```{r}
MA_state_plane <- "+proj=lcc +lat_1=42.68333333333333 +lat_2=41.71666666666667 +lat_0=41 +lon_0=-71.5 +x_0=200000.0001016002 +y_0=750000 +ellps=GRS80 +datum=NAD83 +to_meter=0.3048006096012192 +no_defs"

ggplot(Suffolk_Black) +
  annotation_map_tile(zoomin = 0, progress = "none", type = "stamenbw") +
  geom_sf(color = NA, aes(fill = pct_Black), alpha = 0.7) +
  coord_sf(crs = MA_state_plane) +
  scale_fill_continuous(low="cornsilk", high="darkgreen", 
                       na.value=NA,
                       name = "Percent of population\nidentifying as Black alone",
                       breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1),
                       labels = c("0", "20%", "40%", "60%", "80%", "100%")) +
  theme_void()
```

### Create Interactive Map
```{r}
MA_Black_palette <- colorNumeric(c("cornsilk", "darkgreen"), Suffolk_Black$pct_Black)

Black_map1 <- leaflet(Suffolk_Black) %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(stroke = FALSE, fillOpacity = 0.7,
    color = ~MA_Black_palette(pct_Black)) %>%
  addLegend("bottomright", pal = MA_Black_palette, values = ~pct_Black,
    title = "Percent of population<br/>identifying as Black alone",
    labFormat = labelFormat(suffix = "%",
                            transform = function(x) 100 * x),
    opacity = 1)

Black_map1
```

### Add Generic Pop Ups and Labels
```{r}
Black_map2 <- leaflet(Suffolk_Black) %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(color = ~MA_Black_palette(pct_Black), stroke = FALSE, fillOpacity = 0.7,
              highlightOptions = highlightOptions(fillColor = "darkorange", fillOpacity = 0.9),
              label = "This is a label",
              popup = "This is a popup") %>%
    addLegend("bottomright", pal = MA_Black_palette, values = ~pct_Black,
    title = "Percent of population<br/>identifying as Black alone",
    labFormat = labelFormat(suffix = "%",
                            transform = function(x) 100 * x),
    opacity = 1)

Black_map2
```

### Add Specific Pop Ups and Labels
```{r}
Black_map3 <- leaflet(Suffolk_Black) %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(color = ~MA_Black_palette(pct_Black), stroke = FALSE, fillOpacity = 0.7,
              highlightOptions = highlightOptions(fillColor = "darkorange", fillOpacity = 0.9),
              label = Suffolk_Black$NAME,
              popup = paste("Total population: ", Suffolk_Black$tot_pop, "<br/>",
                            "Black population: ", Suffolk_Black$bl_pop, " (", 
                            round(Suffolk_Black$pct_Black * 100, 1), "%)", sep = "")) %>%
    addLegend("bottomright", pal = MA_Black_palette, values = ~pct_Black,
    title = "Percent of population<br/>identifying as Black alone",
    labFormat = labelFormat(suffix = "%",
                            transform = function(x) 100 * x),
    opacity = 1)

Black_map3
```

### Save Interactive Map as Standalone HTML File
```{r}
saveWidget(Black_map3, file="inter_black.html")
```

### Load ACS Data, Modify Data
```{r}
transit_Suffolk <- get_acs(geography = "tract", county = "Suffolk", state = "MA", 
                           year = 2019, survey = "acs5",
                           variables = c(tot_wrkrs = "B08301_001", pt_wrkrs = "B08301_010"),
                           output = "wide", geometry = TRUE) 

transit_Suffolk <- transit_Suffolk %>%
  select(-tot_wrkrsM, -pt_wrkrsM) %>%
  filter(tot_wrkrsE > 0) %>%
  mutate(pct_transit = round(100 * pt_wrkrsE / tot_wrkrsE, 1))
```

### Plot ACS Data on Interactive Map
```{r}
transit_palette <- colorNumeric(c("pink", "lightblue"),
                                transit_Suffolk$pct_transit)

transit_map <- leaflet(transit_Suffolk) %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(fillColor = ~transit_palette(pct_transit), weight = 1, color = "gray", fillOpacity = 0.7,
              highlightOptions = highlightOptions(fillColor = "yellow", fillOpacity = 0.9),
              label = transit_Suffolk$NAME,
              popup = paste("Total workers: ", transit_Suffolk$tot_wrkrsE, "<br/>",
                            "Transit commuters: ", transit_Suffolk$pt_wrkrsE, " (", 
                            transit_Suffolk$pct_transit, "%)", sep = "")) %>%
    addLegend("bottomright", pal = transit_palette, values = ~pct_transit,
    title = "Percent of workers<br/>communting by transit",
    labFormat = labelFormat(suffix = "%"),
    opacity = 1)

transit_map
```

### Load Transit Stop Data
```{r}
MBTA_url <- feedlist[feedlist$t == "MBTA GTFS",]$url_d

MBTA <- read_gtfs(MBTA_url)

transit_stops <- MBTA$stops

transit_stations <- transit_stops %>%
  filter(location_type == 1)
```

### Convert Point Data to/from CSV to Create Data Frame with Separate Columns for Geography
```{r}
station_sf <- st_as_sf(transit_stations, 
                          coords = c("stop_lon", "stop_lat"), 
                          crs = "WGS84")

st_write(station_sf, "MBTA_stations.csv", layer_options = "GEOMETRY=AS_XY", append = FALSE)

stations_2 <- read_csv("MBTA_stations.csv")
```

### Displaying Points as Markers
```{r}
transit_map2 <- transit_map %>%
  addMarkers(lng = transit_stations$stop_lon,
             lat = transit_stations$stop_lat,
             popup = transit_stations$stop_name)

transit_map2
```

### Set Map Extents and User Bounds
```{r}
limits <- st_bbox(transit_Suffolk)

transit_map3 <- transit_map2 %>%
   fitBounds( lng1 = as.numeric(limits[1]),
                lat1 = as.numeric(limits[2]),
                lng2 = as.numeric(limits[3]),
                lat2 = as.numeric(limits[4])) %>%
   setMaxBounds( lng1 = as.numeric(limits[1]),
                lat1 = as.numeric(limits[2]),
                lng2 = as.numeric(limits[3]),
                lat2 = as.numeric(limits[4])) 

transit_map3
```

### Display Points as Circle Markets
```{r}
transit_map4 <- transit_map %>%
  addCircleMarkers(stroke = FALSE, color = "black", fillOpacity = 1, radius = 3,
                   lng = transit_stations$stop_lon,
                   lat = transit_stations$stop_lat,
                   popup = transit_stations$stop_name) %>%
   fitBounds( lng1 = as.numeric(limits[1]),
                lat1 = as.numeric(limits[2]),
                lng2 = as.numeric(limits[3]),
                lat2 = as.numeric(limits[4])) %>%
   setMaxBounds( lng1 = as.numeric(limits[1]),
                lat1 = as.numeric(limits[2]),
                lng2 = as.numeric(limits[3]),
                lat2 = as.numeric(limits[4])) 

transit_map4
```